The Magnetorotational Instability Prefers Three Dimensions

### 2019-09-20 paper/paper.tex
Modifications to author contributions.
two tiny changes. otherwise all good

We can eliminate the first-order term via
\Beq
\hat{V}_{x}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
\Eeq
\Beq
-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}
\Eeq
\Beq
\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
   \left(\beta '(x)-2 \gamma (x)\right)+\beta
   (x)^2}{4 \alpha (x)^2}.
\Eeq
Finally, by putting everything together 
\Beq
\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
   _S^2}{\left(\omega _A^2-\omega
   _S^2\right){}^2} + k^{2} + \ell^{2}.
\Eeq
If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$, the denominator of $\Phi(x)$ cannot vanish. 
We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
\Beq
\psi(x=\pm d/2) = 0.
\Eeq
To do so, we solve perturbatively near the critical values for the 2D instability. We rescale each parameter in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell,
\Eeq 
keeping in mind
\Beq
q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}.
\Eeq
Expanding equation~(\ref{Schr-eq}) to $\mathcal{O}(\eps^{2})$, we arrive at
\Beq
\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 q \left(k^2 R+q
   \ell ^2\right) - B_{0}^4 k^4 +f^2 (q+1) (\omega -f q x \ell
   )^2  }{B_{0}^4 k^2}.
\Eeq
The leading order balance is
\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0  \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d),
\Eeq
and next order is
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0.
\Eeq
The solvability condition for $\psi_{2}(x)$ is 
\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq
This yields our final result
\begin{equation}\label{eq:asymp}
\sigma^{2}=\frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
\end{equation}
where
\begin{equation}
\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4},
\end{equation}
and we have switched back to $\sigma$ for time dependence.
The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_{y}\neq0$ and ultraviolet divergence in the absence of higher-order effects.

@@

\aucontribute{J\@.S\@.O and G\@.M.\@V\@. led the project.
G\@.M\@.V performed the asymptotic calculation.
M.B.\ ran the bulk of the eigenvalue calculations and prepared several figures.
A\@.S\@. performed preliminary eigenvalue calculations.
J\@.S\@.O, G\@.M.V\@., K\@.J\@.B., D\@.L\@., and B.P\@.B developed \emph{Dedalus}, which is used for all calculations in this paper.
All authors contributed to the writing of the manuscript.
}

>>>>

We can eliminate the first-order term via
\Beq
\hat{V}_{x}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
\Eeq
\Beq
-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}
\Eeq
\Beq
\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
   \left(\beta '(x)-2 \gamma (x)\right)+\beta
   (x)^2}{4 \alpha (x)^2}.
\Eeq
Finally, by putting everything together 
\Beq
\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
   _S^2}{\left(\omega _A^2-\omega
   _S^2\right){}^2} + k^{2} + \ell^{2}.
\Eeq
If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$, the denominator of $\Phi(x)$ cannot vanish. 
We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
\Beq
\psi(x=\pm d/2) = 0.
\Eeq
To do so, we solve perturbatively near the critical values for the 2D instability. We rescale each parameter in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell,
\Eeq 
keeping in mind
\Beq
q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}.
\Eeq
Expanding equation~(\ref{Schr-eq}) to $\mathcal{O}(\eps^{2})$, we arrive at
\Beq
\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 q \left(k^2 R+q
   \ell ^2\right) - B_{0}^4 k^4 +f^2 (q+1) (\omega -f q x \ell
   )^2  }{B_{0}^4 k^2}.
\Eeq
The leading order balance is
\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0  \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d),
\Eeq
and next order is
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0.
\Eeq
The solvability condition for $\psi_{2}(x)$ is 
\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq
This yields our final result
\begin{equation}\label{eq:asymp}
\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
\end{equation}
where
\begin{equation}
\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4},
\end{equation}
and we have switched back to $\sigma$ for time dependence.
The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_{y}\neq0$ and ultraviolet divergence in the absence of higher-order effects.

@@

\aucontribute{J\@.S\@.O and G\@.M.\@V\@. led the project.
G\@.M\@.V performed the asymptotic calculation.
M.B.\ ran the bulk of the eigenvalue calculations and prepared several preliminary figures.
A\@.S\@. performed preliminary eigenvalue calculations which led to the fiducial parameter choices used here.
J\@.S\@.O, G\@.M.V\@., K\@.J\@.B., D\@.L\@., and B.P\@.B developed \emph{Dedalus}, which is used for all calculations in this paper.
All authors contributed to the writing of the manuscript and physical interpretation of the results.
}
